Long-form recording of infant body position in the home using wearable inertial sensors

### 2023-08-14 manuscript.tex
Working on revision

\abstract{%
Long form audio recordings have had a transformational effect on the study of infant language acquisition by using mobile, unobtrusive devices to gather full-day, real-time data that can be automatically scored. How can we produce similar data in service of measuring infants' everyday motor behaviors, such as body position? The aim of the current study was to validate long form recordings of infant position (supine, prone, sitting, upright, held by caregiver) based on machine learning classification of data from inertial sensors worn on infants' ankles and thighs. Using over 100 hours of video recordings synchronized with inertial sensor data from infants in their homes, we demonstrate that body position classifications are sufficiently accurate to measure infant behavior. Moreover, classification remained accurate when predicting behavior later in the session when infants and caregivers were unsupervised and went about their normal activities, showing that the method can handle the challenge of measuring unconstrained, natural activity. Next, we show that the inertial sensing method has convergent validity by replicating age differences in body position found using other methods with full-day data captured from inertial sensors. We end the paper with a discussion of the novel opportunities that long form motor recordings afford for understanding infant learning and development.
}

@@

In this paper, we present an inertial sensing method to classify infants' full-day, real-time body position. Our method takes inspiration from a more mature technology: Long-form audio recordings of infants' language experiences. We begin by identifying the key features of wearable audio recorders that should be replicated in long-form recordings of motor behavior. Next, we review the current state-of-the-art in measuring infant motor behavior---video and survey data---and their limitations in capturing real-time, full-day behavior. Finally, we discuss the advantages of using inertial sensing to classify motor behavior. Despite promising past results in brief, supervised sessions \autocite{FranchakScott2021,AiraksinenRasanen2020,AiraksinenGallen2022}, the current investigation takes a needed step forward by testing accuracy over long, unsupervised recordings in what we term a \emph{distal comparison}.

@@

But automatic classification can only replace human annotation if it is sufficiently accurate. An independent assessment of the LENA® algorithms found mixed results about classification accuracy. Correlations between human transcribed counts of adult words and child vocalizations against LENA®'s automatic counts were strong, \emph{r} = .698 and \emph{r} = .649, respectively \autocite{CristiaLavechin2020}. For other metrics, such as the number of ``conversational turns'' between the child and communicative partners, agreement was poor (\emph{r} = .364). Thus, for some use cases (and for some metrics), long form audio recordings provide a mobile, unobtrusive way to automatically score real-time data over a full day.

@@

Measuring infant movement with inertial movement units (IMUs) is a promising avenue for long form recordings of motor behavior in the home \autocite{CliffReilly2009,Barbaro2019,BruijnsTruelove2020,LoboHall2019}. Lightweight sensors (10-30 g) can be embedded in garments to make recordings fully \emph{mobile}, and they are \emph{unobtrusive} because they do not require a researcher to follow with a camcorder. Many commercially-available IMUs have \textgreater{} 12 hour battery life with onboard storage to record \emph{real-time}, \emph{full-day} motion data at a high sampling rate (e.g., 50-100 Hz).

@@

Modern approaches to human activity recognition have used machine learning to classify activity categories based on features derived from IMU data in adults \autocite{Preece2009,Arif2015}, children \autocite{NamPark2013,RenDing2016,StewartNarayanan2018}, and infants \autocite{YaoPlotz2019,FranchakScott2021,AiraksinenRasanen2020}. Three prior investigations have used different machine learning techniques to categorize infant body position from IMUs towards the goal of collecting full-day data. \textcite{AiraksinenRasanen2020} tested 4- to 8-month-olds in a laboratory visit with a 4-sensor array (one on each thigh and one below each shoulder), and found 95\% accuracy in distinguishing between body position categories that crawling infants could perform on the floor (excluding times that infants were held by caregivers). Using a wider age range of 6-18 months, \textcite{FranchakScott2021} found 98\% accuracy (\emph{kappa} = 95\%) in a laboratory validation study with a 3-sensor array (ankle, knee, thigh on a single leg) in categorizing body position that included infants who could both crawl and walk and also included a category for caregiver holding. Most recently, \textcite{AiraksinenGallen2022} conducted a validation study of body position classification in either a home or clinic testing 4- to 19-month-olds with their 4-sensor system, refining their previous method to detect moments that infants were carried by caregivers. Classification accuracy did not vary between lab and home settings, and was generally high (95\%, \emph{kappa} = .93). Although all three studies yielded promising classification accuracy, accuracy was assessed in brief (15-60 minute) sessions supervised by a researcher, leaving the open question of how well body position classification will scale to testing across an entire day of natural home life.

@@

Infants were recruited in one of two age groups: \emph{Younger} infants were between 4 and 7 months and \emph{older} infants were between 11 and 14 months. There were 8 infants in the 4-7 month group and 14 in the 11-14 month group. Ten infants were female and 12 were male. Families were recruited through social media advertisements and from community events in Southern California. Most infants were reported to be either Hispanic and White (\emph{n} = 8) or Non-Hispanic and White (\emph{n} = 7). Families were compensated \$30 for every home recording session they completed. The BLINDED Institutional Review Board reviewed and approved all procedures associated with the study. All caregivers gave their informed consent before the start of the study.

Most participants were tested in a single session (\emph{n} = 15), but 7 participants contributed between 2-4 sessions as part of an ongoing longitudinal study (three from the younger group and four from the older group). Although including more data from some participants could over-represent their characteristics in the model, we reasoned that this drawback was outweighed by having more available data to use for training. Only 1 session was excluded due to a technical error---one of the four IMU sensors failed to record, resulting in an unusable set of data for classification. Across the two age groups, we report data on a total of 34 sessions, with 14 sessions from younger infants and 20 sessions from older infants. Across sessions, younger infants' age ranged from 3.8 to 7.2 months (\emph{M} = 5.2) and older infants' age ranged from 10.7 to 14.2 months (\emph{M} = 11.7). The total number of recording sessions (34) exceeded the number of sessions employed in comparable past work: 10 in \textcite{NamPark2013}, 15 in \textcite{FranchakScott2021}, 22 in \textcite{AiraksinenRasanen2020}, and 33 in \textcite{YaoPlotz2019}.

@@

Because the camera only had the battery life to record for \textasciitilde3 hours (automatically split into two 90-minute video files), this divided the day into different periods for analysis. As seen in the bottom of Figure \ref{fig:exemplar-timeline}), the \emph{video period} comprised the first three hours of recording starting from the researcher's arrival when they turned on the camera. The first 90-minute video file, termed the \emph{proximal comparison}, contained the activities during the guided call followed by a period during which infants and caregivers resumed their normal activities. Because this video contained the synchronization point, the data in this period had high temporal synchrony between IMU and video data. Synchronization errors were estimated to be less than 30-60 ms (1-2 video frames). The second video file comprised the \emph{distal comparison}. This video recorded the next 90 minutes of natural activity. However, because of a limitation in the camera, there was a variable gap of \textasciitilde40 s between the two videos, so synchronization in the distal comparison video was coarser, with estimated temporal offsets of \textasciitilde5 s in either direction.

@@

The proximal and distal comparison videos for each participant were annotated by trained human coders to identify infant \emph{body position} into one of 5 mutually-exclusive categories following the definitions in prior work \autocite{FranchakScott2021}: supine, prone, sitting, upright, or held by caregiver. All coding was done using Datavyu software (datavyu.org). A Databrary repository contains the entire video recording, coding files, and raw IMU data for a single participant (\url{https://nyu.databrary.org/volume/1580}).

@@

\begin{tabular}{lllllll}
\toprule
 & \multicolumn{3}{c}{Group} & \multicolumn{3}{c}{Individual} \\
\cmidrule(r){2-4} \cmidrule(r){5-7}
Sensors & Median & Mean & SD & Median & Mean & SD\\
\midrule
All & 0.894 & 0.846 & 0.131 & 0.933 & 0.916 & 0.072\\ \midrule
Left Hip/Ankle & 0.891 & 0.824 & 0.177 & 0.915 & 0.901 & 0.071\\
Right Thigh/Ankle & 0.842 & 0.800 & 0.167 & 0.897 & 0.898 & 0.086\\
Both Thighs & 0.859 & 0.791 & 0.195 & 0.907 & 0.898 & 0.072\\
Both Ankles & 0.797 & 0.719 & 0.196 & 0.864 & 0.844 & 0.108\\ \midrule
Left Thigh & 0.842 & 0.762 & 0.208 & 0.889 & 0.872 & 0.077\\
Left Ankle & 0.717 & 0.667 & 0.208 & 0.851 & 0.825 & 0.117\\
Right Thigh & 0.846 & 0.772 & 0.178 & 0.896 & 0.876 & 0.092\\
Right Ankle & 0.704 & 0.683 & 0.160 & 0.830 & 0.819 & 0.114\\
\bottomrule
\end{tabular}

@@

The first set of results showed that group and individual models trained from data during the proximal comparison period were within the proximal period. In the next analysis, we examine long-term performance by testing how accurately models trained from the proximal period could predict body position during the distal comparison period. A single group model was created using all sessions' proximal period training data (rather than group models leaving out a single session); the same individual models were used. Distal videos had only coarse temporal synchrony with motion recordings which precluded calculating accuracy based on the proportion of matching events. Instead, we summed the amount of time infants were predicted to be in each of the 5 body position categories from the model and compared that to the summed time for the body positions based on human coding \autocite{YaoPlotz2019,FranchakScott2021}.

@@

Figure \ref{fig:part2overall} and Table \ref{tab:pt2overalltable} summarize the overall agreement during the distal comparison period. For each session, we calculated the actual time spent in each body position (out of times the infant was visible on camera and awake) using human annotated body position (x-axis on Figure \ref{fig:part2overall}). Predicted time was calculated the same way for group and individual model predictions, omitting the off-camera and nap periods to make a direct comparison. Overall---across participants and across body position classes---agreement was strong: The correlation between group model predictions and human-coded time was \emph{r} = 0.80, and the correlation between individual model predictions and human-coded time was \emph{r} = 0.91. As in the proximal comparison, agreement varied somewhat between body positions; in particular, agreement for held was poor. Unlike in the proximal period, some body positions were better predicted by group models and others by individual models.

@@

The challenge increased in the distal comparison period. Since sessions were scheduled at convenient times for the infant and caregiver, the transition from the proximal to distal period increased the odds that infants might need a nap, a snack/meal, or engage in a less typical activity (e.g., watching TV). Indeed, the distal comparison in Figure \ref{fig:exemplar-timeline} contains a long period (almost 50 minutes) where the infant sat in a high chair eating lunch. Regardless of these challenges, accuracy in the distal comparison continued to be strong. Barring two outliers (which we will discuss further), agreement was high for group (\emph{r} = 0.92) and individual models (\emph{r} = 0.94). Even at a finer timescale, most errors within 10-minute bins were less than 1 minute in total for all five body positions. For comparison, the reliability of human coders on the body position code was \textasciitilde90\%, putting model-predicted accuracy on a similar level to human coders.

@@

Long-form recordings of motor behavior bring about new research possibilities. Although past work using video \autocite{Karasik2011,ChenSchneider2022,HerzbergFletcher2021} produced real-time data, such data were limited to a relatively small part of the infants' day. And although survey methods \autocite{Survey} could capture moments scattered throughout the day, they do not produce real-time data. The combination of real-time, full-day data about infant motor behavior is unprecedented and offers new opportunities for understanding infants' everyday experiences. Collecting dense data over the entire variety of daily experiences helps to more accurately measure infants' experiences in aggregate (as in Figure \ref{fig:age}) without biasing results from a particular type of activity (e.g., play). Long-form recordings also have the potential to measure clinically-relevant outcomes for infants with motor delay or other pediatric concerns.

@@

We acknowledge several limitations in the current approach. First, a stronger test of full-day accuracy would be to compare accuracy at the start of the recording period to accuracy at the end, rather than using the distal comparison video that followed the first 90 minutes of the study. As we described in the procedure section, synchronizing video cameras and inertial sensors required capturing a synchronization point on video. We completed this procedure prior to giving equipment to caregivers so that they were not responsible for synchronization; in fact, caregivers never (purposefully) touched a button on the video camera. An end-of-day video would require a lot of effort and compliance on the part of caregivers, and any mistakes in the procedure would lead to misaligned and unusable data. Given that most recordings ended at infants' bedtimes, having an experimenter visit families in the evening to video record would be intrusive.

Second, we relied on caregiver logs of infants' naps and times when sensors were removed to separate usable data (times when infants wore the sensors while awake) from unusable data. Future work should aim to further assess caregivers' perceptions about the usability and comfort of the garment. Although infants wore the sensor garment throughout all of the desired times, we did not collect independent data (aside from caregivers' logs) to verify wear time. Most caregivers were diligent about completing logs, however, there were a few cases that caregivers may have failed to report naps (younger infants \#1 and \#4). In the future, algorithms can be developed to automatically identify periods of sleep and times when sensors are off the body. Such algorithms exist for adults using wrist-worn sensors, however, we did not use them because they have not been validated for infant participants wearing sensors on thighs/ankles. Differences in infants' positioning while asleep (held in caregivers arms, laying in cribs, reclined in strollers and carseats) might make classifying sleep more difficult compared with adults based on movement.

Third, although the goal was to collect fully naturalistic behavior, our recording protocol led to some changes that might have affected (or missed) some behaviors. We opted to restrict recording to times when participants were in the home. Inertial sensors can travel with participants, and there is no doubt that infants could have worn the sensors out of the home on errands. However, we do not yet have training data to detect periods of transportation (e.g., moving in a car or in a stroller) that might add noise and/or lead to misclassification of body positions. It is also unknown whether participants might have behaved differently had they not worn the sensors. The added bulk in the garment from the sensors (including the LENA® recorder) might have made some positions more uncomfortable, such as lying prone with the LENA® recorder worn on the chest, and/or influenced how caregivers chose to position infants throughout the day.

>>>>

\abstract{%
Long-form audio recordings have had a transformational effect on the study of infant language acquisition by using mobile, unobtrusive devices to gather full-day, real-time data that can be automatically scored. How can we produce similar data in service of measuring infants' everyday motor behaviors, such as body position? The aim of the current study was to validate long-form recordings of infant position (supine, prone, sitting, upright, held by caregiver) based on machine learning classification of data from inertial sensors worn on infants' ankles and thighs. Using over 100 hours of video recordings synchronized with inertial sensor data from infants in their homes, we demonstrate that body position classifications are sufficiently accurate to measure infant behavior. Moreover, classification remained accurate when predicting behavior later in the session when infants and caregivers were unsupervised and went about their normal activities, showing that the method can handle the challenge of measuring unconstrained, natural activity. Next, we show that the inertial sensing method has convergent validity by replicating age differences in body position found using other methods with full-day data captured from inertial sensors. We end the paper with a discussion of the novel opportunities that long-form motor recordings afford for understanding infant learning and development.
}

@@

In this paper, we present an inertial sensing method to classify infants' full-day, real-time body position. Our method takes inspiration from a more mature technology: Long-form audio recordings of infants' language experiences. We begin by identifying the key features of wearable audio recorders that should be replicated in long-form recordings of motor behavior. Next, we review the current state-of-the-art in measuring infant motor behavior---video and survey data---and their limitations in capturing real-time, full-day behavior. Finally, we discuss the advantages of using inertial sensing to classify motor behavior. Despite promising past results in brief, supervised sessions \autocite{FranchakScott2021,AiraksinenRasanen2020,GreenspanCunha2021,AiraksinenGallen2022}, the current investigation takes a needed step forward by testing accuracy over long, unsupervised recordings in what we term a \emph{distal comparison}.

@@

But automatic classification can only replace human annotation if it is sufficiently accurate. An independent assessment of the LENA® algorithms found mixed results about classification accuracy. Correlations between human transcribed counts of adult words and child vocalizations against LENA®'s automatic counts were strong, \emph{r} = .698 and \emph{r} = .649, respectively \autocite{CristiaLavechin2020}. For other metrics, such as the number of ``conversational turns'' between the child and communicative partners, agreement was poor (\emph{r} = .364). Thus, for some use cases (and for some metrics), long-form audio recordings provide a mobile, unobtrusive way to automatically score real-time data over a full day.

@@

Measuring infant movement with inertial movement units (IMUs) is a promising avenue for long-form recordings of motor behavior in the home \autocite{CliffReilly2009,Barbaro2019,BruijnsTruelove2020,LoboHall2019}. Lightweight sensors (10-30 g) can be embedded in garments to make recordings fully \emph{mobile}, and they are \emph{unobtrusive} because they do not require a researcher to follow with a camcorder. Many commercially-available IMUs have \textgreater{} 12 hour battery life with onboard storage to record \emph{real-time}, \emph{full-day} motion data at a high sampling rate (e.g., 50-100 Hz).

@@

Modern approaches to human activity recognition have used machine learning to classify activity categories based on features derived from IMU data in adults \autocite{Preece2009,Arif2015}, children \autocite{NamPark2013,RenDing2016,StewartNarayanan2018}, and infants \autocite{YaoPlotz2019,FranchakScott2021,AiraksinenRasanen2020}. Three prior investigations have used different machine learning techniques to categorize infant body position from multiple IMUs towards the goal of collecting full-day data. \textcite{AiraksinenRasanen2020} tested 4- to 8-month-olds in a laboratory visit with a 4-sensor array (one on each thigh and one below each shoulder), and found 95\% accuracy in distinguishing between body position categories that crawling infants could perform on the floor (excluding times that infants were held by caregivers). Using a wider age range of 6-18 months, \textcite{FranchakScott2021} found 98\% accuracy (\emph{kappa} = 95\%) in a laboratory validation study with a 3-sensor array (ankle, knee, thigh on a single leg) in categorizing body position that included infants who could both crawl and walk and also included a category for caregiver holding. Most recently, \textcite{AiraksinenGallen2022} conducted a validation study of body position classification in either a home or clinic testing 4- to 19-month-olds with a 4-sensor system, refining their previous method to detect moments that infants were carried by caregivers. Classification accuracy did not vary between lab and home settings, and was generally high (95\%, \emph{kappa} = .93). In contrast to the three machine learning studies that used 3-4 sensors \autocite{AiraksinenRasanen2020,FranchakScott2021,AiraksinenGallen2022}, \textcite{GreenspanCunha2021} used orientation from a single hip-worn sensor to measure body position with a high degree of accuracy (\emph{kappa} = .84). Although all four studies yielded promising classification accuracy, accuracy was assessed in brief (15-60 minute) sessions supervised by a researcher, leaving the open question of how well body position classification will scale to testing across an entire day of natural home life.

@@

Infants were recruited in one of two age groups: \emph{Younger} infants were between 4 and 7 months and \emph{older} infants were between 11 and 14 months. There were 8 infants in the 4-7 month group and 14 in the 11-14 month group. Ten infants were female and 12 were male. Families were recruited through social media advertisements and from community events in Southern California. Most infants were reported to be either Hispanic and White (\emph{n} = 8) or Non-Hispanic and White (\emph{n} = 7). Families were compensated \$30 for every home recording session they completed. The University of California, Riverside Institutional Review Board reviewed and approved all procedures associated with the study. All caregivers gave their informed consent before the start of the study.

Most participants were tested in a single session (\emph{n} = 15), but 7 participants contributed between 2-4 sessions as part of an ongoing longitudinal study (3 from the younger group and 4 from the older group). Although including more data from some participants could over-represent their characteristics in the model, we reasoned that this drawback was outweighed by having more available data to use for training. Only 1 session was excluded due to a technical error---one of the four IMU sensors failed to record, resulting in an unusable set of data for classification. Across the two age groups, we report data on a total of 34 sessions, with 14 sessions from younger infants and 20 sessions from older infants. Across sessions, younger infants' age ranged from 3.8 to 7.2 months (\emph{M} = 5.2) and older infants' age ranged from 10.7 to 14.2 months (\emph{M} = 11.7). The total number of recording sessions (34) exceeded the number of sessions employed in comparable past work: 10 in \textcite{NamPark2013}, 15 in \textcite{FranchakScott2021}, 22 in \textcite{AiraksinenRasanen2020}, 23 in \textcite{GreenspanCunha2021}, and 33 in \textcite{YaoPlotz2019}.

@@

Because the camera only had the battery life to record for \textasciitilde3 hours (automatically split into two 90-minute video files), this divided the day into different periods for analysis. As seen in the bottom of Figure \ref{fig:exemplar-timeline}, the \emph{video period} comprised the first three hours of recording starting from the researcher's arrival when they turned on the camera. The first 90-minute video file, termed the \emph{proximal comparison}, contained the activities during the guided call followed by a period during which infants and caregivers resumed their normal activities. Because this video contained the synchronization point, the data in this period had high temporal synchrony between IMU and video data. Synchronization errors were estimated to be less than 30-60 ms (1-2 video frames). The second video file comprised the \emph{distal comparison}. This video recorded the next 90 minutes of natural activity. However, because of a limitation in the camera, there was a variable gap of \textasciitilde40 s between the two videos, so synchronization in the distal comparison video was coarser, with estimated temporal offsets of \textasciitilde5 s in either direction.

@@

The proximal and distal comparison videos for each participant were annotated by trained human coders to classify infant \emph{body position} into one of 5 mutually-exclusive categories following the definitions in prior work \autocite{FranchakScott2021}: supine, prone, sitting, upright, or held by caregiver. All coding was done using Datavyu software (datavyu.org). A Databrary repository contains the entire video recording, coding files, and raw IMU data for a single participant (\url{https://nyu.databrary.org/volume/1580}).

@@

\begin{tabular}{lllllll}
\toprule
 & \multicolumn{3}{c}{Group} & \multicolumn{3}{c}{Individual} \\
\cmidrule(r){2-4} \cmidrule(r){5-7}
Sensors & Median & Mean & SD & Median & Mean & SD\\
\midrule
All & 0.894 & 0.846 & 0.131 & 0.933 & 0.916 & 0.072\\ \midrule
Left Thigh/Ankle & 0.891 & 0.824 & 0.177 & 0.915 & 0.901 & 0.071\\
Right Thigh/Ankle & 0.842 & 0.800 & 0.167 & 0.897 & 0.898 & 0.086\\
Both Thighs & 0.859 & 0.791 & 0.195 & 0.907 & 0.898 & 0.072\\
Both Ankles & 0.797 & 0.719 & 0.196 & 0.864 & 0.844 & 0.108\\ \midrule
Left Thigh & 0.842 & 0.762 & 0.208 & 0.889 & 0.872 & 0.077\\
Left Ankle & 0.717 & 0.667 & 0.208 & 0.851 & 0.825 & 0.117\\
Right Thigh & 0.846 & 0.772 & 0.178 & 0.896 & 0.876 & 0.092\\
Right Ankle & 0.704 & 0.683 & 0.160 & 0.830 & 0.819 & 0.114\\
\bottomrule
\end{tabular}

@@

The first set of results showed that group and individual models trained from data during the proximal comparison period were accurate during the proximal period. In the next analysis, we examine long-term performance by testing how accurately models trained from the proximal period could predict body position during the distal comparison period. A single group model was created using all sessions' proximal period training data (rather than group models leaving out a single session); the same individual models were used. Distal videos had only coarse temporal synchrony with motion recordings which precluded calculating accuracy based on the proportion of matching events. Instead, we summed the amount of time infants were predicted to be in each of the 5 body position categories from the model and compared that to the summed time for the body positions based on human coding \autocite{YaoPlotz2019,FranchakScott2021}.

@@

Figure \ref{fig:part2overall} and Table \ref{tab:pt2overalltable} summarize the overall agreement during the distal comparison period. For each session, we calculated the actual time spent in each body position (out of times the infant was visible on camera and awake) using human annotated body position (x-axis on Figure \ref{fig:part2overall}). Predicted time was calculated the same way for group and individual model predictions, omitting the off-camera and nap periods to make a direct comparison. Agreement was strong across participants and across body position classes: The correlation between group model predictions and human-coded time was \emph{r} = 0.80, and the correlation between individual model predictions and human-coded time was \emph{r} = 0.91. As in the proximal comparison, agreement varied somewhat between body positions; in particular, agreement for held was poor. Unlike in the proximal period, some body positions were better predicted by group models and others by individual models.

@@

The challenge increased in the distal comparison period. Since sessions were scheduled at convenient times for the infant and caregiver, the transition from the proximal to distal period increased the odds that infants might need a nap, eat a snack/meal, or engage in a less typical activity (e.g., watching TV). Indeed, the distal comparison in Figure \ref{fig:exemplar-timeline} contains a long period (almost 50 minutes) where the infant sat in a high chair eating lunch. Regardless of these challenges, accuracy in the distal comparison continued to be strong. Barring two outliers (which we will discuss further), agreement was high for group (\emph{r} = 0.92) and individual models (\emph{r} = 0.94). Even at a finer timescale, most errors within 10-minute bins were less than 1 minute in total for all five body positions. For comparison, the reliability of human coders on the body position code was \textasciitilde90\%, putting model-predicted accuracy on a similar level to human coders.

@@

Long-form recordings of motor behavior bring about new research possibilities. Although past work using video \autocite{Karasik2011,ChenSchneider2022,HerzbergFletcher2021} produced real-time data, such data were limited to a relatively small part of the day. And although survey methods \autocite{Survey} could capture moments scattered throughout the day, they do not produce real-time data. The combination of real-time, full-day data about infant motor behavior is unprecedented and offers new opportunities for understanding infants' everyday experiences. Collecting dense data over the entire variety of daily experiences helps to more accurately measure infants' experiences in aggregate (as in Figure \ref{fig:age}) without biasing results from a particular type of activity (e.g., play). Long-form recordings also have the potential to measure clinically-relevant outcomes for infants with motor delay or other pediatric concerns.

@@

We acknowledge several limitations in the current approach. First, despite the large amount of data collected per infant, the sample of infants was small (22 unique participants). Since the study combined data from participants who completed single sessions and participants who completed multiple sessions, the seven infants with multiple sessions are over-represented in the dataset. A larger, more representative sample would be needed to determine whether the method would generalize to the broader population.

Second, a stronger test of full-day accuracy would be to compare accuracy at the start of the recording period to accuracy at the end, rather than using the distal comparison video that followed the first 90 minutes of the study. As we described in the procedure section, synchronizing video cameras and inertial sensors required capturing a synchronization point on video. We completed this procedure prior to giving equipment to caregivers so that they were not responsible for synchronization; in fact, caregivers never (purposefully) touched a button on the video camera. An end-of-day video would require a lot of effort and compliance on the part of caregivers, and any mistakes in the procedure would lead to misaligned and unusable data. Given that most recordings ended at infants' bedtimes, having an experimenter visit families in the evening to video record would be intrusive.

Third, we relied on caregiver logs of infants' naps and times when sensors were removed to separate usable data (times when infants wore the sensors while awake) from unusable data. Future work should aim to further assess caregivers' perceptions about the usability and comfort of the garment. Although infants wore the sensor garment throughout all of the desired times, we did not collect independent data (aside from caregivers' logs) to verify wear time. Most caregivers were diligent about completing logs, however, there were a few cases that caregivers may have failed to report naps (younger infants \#1 and \#4). In the future, algorithms can be developed to automatically identify periods of sleep and times when sensors are off the body. Such algorithms exist for adults using wrist-worn sensors, however, we did not use them because they have not been validated for infant participants wearing sensors on thighs/ankles. Differences in infants' positioning while asleep (held in caregivers arms, laying in cribs, reclined in strollers and carseats) might make classifying sleep more difficult compared with adults based on movement.

Fourth, although the goal was to collect fully naturalistic behavior, our recording protocol led to some changes that might have affected (or missed) some behaviors. We opted to restrict recording to times when participants were in the home. Inertial sensors can travel with participants, and there is no doubt that infants could have worn the sensors out of the home on errands. However, we do not yet have training data to detect periods of transportation (e.g., moving in a car or in a stroller) that might add noise and/or lead to misclassification of body positions. It is also unknown whether participants might have behaved differently had they not worn the sensors. The added bulk in the garment from the sensors (including the LENA® recorder) might have made some positions more uncomfortable, such as lying prone with the LENA® recorder worn on the chest, and/or influenced how caregivers chose to position infants throughout the day.
